Flute and kink instabilities in a dynamically twisted flux tube with anisotropic plasma viscosity

### 2022-01-21 paper.tex
Minor Query R11: carefully mentioned peak linear growth rate r_s

\begin{figure}
  \centering
    \begin{subfigure}{0.49\textwidth}
      \includegraphics[width=\linewidth]{suydam_condition_4.pdf}
      \caption{Suydam condition}
      \label{fig:suydam_condition_4}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.49\textwidth}
      \includegraphics[width=\linewidth]{growth_rate_4.pdf}
      \caption{Linear growth rate}
      \label{fig:growth_rate_4}
    \end{subfigure}
\mycaption{Stability and linear growth rate of the flute instability.}{In panel
\ref{fig:suydam_condition_4}, Suydam's stability criterion
\eqref{eq:suydams_criterion} and is contributing terms   are plotted and in
panel \ref{fig:growth_rate_4} the predicted linear growth
rate~\eqref{eq:fluting_growth_rate2} is plotted. Both plots are produced at
$t=20$ for $\eta=10^{-4}$ and using the anisotropic model. The location of the
peak linear growth rate is also shown.}
\label{fig:stability_and_growth}%
\end{figure}

Prior to the onset of either instability, the flux tube is found to be linearly
unstable to perturbations of the form~\eqref{eq:kink_perturbation} at $t=20$
via Suydam's criterion~\eqref{eq:suydams_criterion}
(Figure~\ref{fig:suydam_condition_4}). The criterion represents a balance
between destabilising pressure gradients and stabilising magnetic shear and in
this case, the shear is so small and the pressure gradient so large that the
tube is unstable over a wide range of radii, for $ 0.02 \lessapprox
r \lessapprox 0.29$. The measure of linear fluting growth rate $\gamma$ is
plotted as a function of $r$ at the same time (Figure~\ref{fig:growth_rate_4}).
The location of the peak growth matches nearly exactly the location of the
resonant surface where the observed perturbation grows
(Figure~\ref{fig:swi-4_pressure_13}) and an estimate of the linear growth rate
can be found by averaging $\gamma$ over $r$, giving a growth rate of $0.86$.

@@

The value of $k$ for each perturbation is calculated as $k
= 2\pi/\tilde{\lambda}$ where $\tilde{\lambda}$ is the wavelength of the
perturbation, measured as the difference between the two peaks closest to $z=0$
(thus minimising the influence of line-tying on the measurement). This gives
a value of $k_{\text{flute}}=23.61$ and $k_{\text{kink}}=4.57$ for both
viscosity models. Hence, the observed most unstable fluting perturbation is
that of the form~\eqref{eq:kink_perturbation} where $m=4$ and $k=23.61$ and the
observed kink instability is that where $m=1$ and $k=4.57$. Using these values,
it is observed that the fluting perturbation exactly resonates with the field,
that is $m B_{\theta}(r)/r + kB_z(r) = 0$, at $r=0.125$
(Figure~\ref{fig:resonant_surface_4}). This is precisely the predicted radius
of peak linear growth (Figure~\ref{fig:growth_rate_4}). At this time the
perturbation is close to resonance, that is $m B_{\theta}(r)/r + kB_z(r)
\approx 0$, over a range of radii from $r=0.125$ to $0.2$.

@@

Table~\ref{tab:kink_fluting_params} summarises the quantitative differences
between the results for the two values of the resistivity $\eta$. All values
are calculated from simulations using the anisotropic model with the exception
of $k_{\text{kink}}$ which is measured from isotropic results due to noise in
the anisotropic case (the value of $k_{\text{kink}}$ appears similar, however).
The results of the isotropic cases are qualitatively similar. The radius of
peak $\gamma$ is calculated at time $t=20$. The fluting wavenumber $k$ and
observed $r_s$ are measured at times just prior to the nonlinear development of
the flute instability, that is at $t=22$ when $\eta=10^{-3}$ and $t=26$ when
$\eta = 10^{-4}$. The kink wavenumber is measured at $t=26$ in both cases.
These times allow fair comparison between measurements.

The longitudinal wavenumber $k_{\text{kink}}$ of the observed kink perturbation
remains similar in both cases since the instability is essentially governed by
the twist injected by the driver which remains the same in both cases. In
contrast, the longitudinal wavenumber $k_{\text{flute}}$ of the observed
fluting perturbation is lower in the $\eta=10^{-3}$ cases. This is due to the
different resonant surface within which the perturbation grows, the location
being dictated by the peak of the linear growth rate. Note that the location of
this peak again matches well the location of the observed resonant surface, as
in the $\eta=10^{-4}$ cases. Similar to the $\eta=10^{-4}$ cases, the peak
growth rate predicted by the linear analysis is the same order of magnitude as
the observed growth rate.

>>>>

\begin{figure}
  \centering
    \begin{subfigure}{0.49\textwidth}
      \includegraphics[width=\linewidth]{suydam_condition_4.pdf}
      \caption{Suydam condition}
      \label{fig:suydam_condition_4}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.49\textwidth}
      \includegraphics[width=\linewidth]{growth_rate_4.pdf}
      \caption{Linear growth rate}
      \label{fig:growth_rate_4}
    \end{subfigure}
\mycaption{Stability and linear growth rate of the flute instability.}{In panel
\ref{fig:suydam_condition_4}, Suydam's stability criterion
\eqref{eq:suydams_criterion} and is contributing terms   are plotted and in
panel \ref{fig:growth_rate_4} the predicted linear growth
rate~\eqref{eq:fluting_growth_rate2} is plotted. Both plots are produced at
$t=20$ for $\eta=10^{-4}$ and using the anisotropic model. The location of the
\rev{peak linear growth rate $r_s$} is also shown.}
\label{fig:stability_and_growth}%
\end{figure}

Prior to the onset of either instability, the flux tube is found to be linearly
unstable to perturbations of the form~\eqref{eq:kink_perturbation} at $t=20$
via Suydam's criterion~\eqref{eq:suydams_criterion} as shown in
Figure~\ref{fig:suydam_condition_4}. The criterion represents a
balance between destabilising pressure gradients and stabilising
magnetic shear and in this case, the shear is so small and the
pressure gradient so large that the tube is unstable over a wide range
of radii, for $ 0.02 \lessapprox r \lessapprox 0.29$.
\rev{Using equation~\eqref{eq:fluting_growth_rate2}, the} linear
fluting  growth rate $\gamma$ is plotted as a function of $r$ at
$t=20$ in Figure~\ref{fig:growth_rate_4}. \rev{At any fixed moment,
the radial dependence of the flute instability growth rate, $\gamma(r)$, is a concave
function and peaks at a certain radius that we denote by $r_s$ in
Figure~\ref{fig:growth_rate_4} and Table \ref{tab:kink_fluting_params}
where we compare theoretical and numerical estimates.}

The location of the peak \rev{fluting growth rate $r_s$} matches
nearly exactly the location of the resonant surface where the observed perturbation grows
(Figure~\ref{fig:swi-4_pressure_13}) and an estimate of the linear growth rate
can be found by averaging $\gamma$ over $r$, giving a growth rate of $0.86$.

@@

The value of $k$ for each perturbation is calculated as $k
= 2\pi/\tilde{\lambda}$ where $\tilde{\lambda}$ is the wavelength of the
perturbation, measured as the difference between the two peaks closest to $z=0$
(thus minimising the influence of line-tying on the measurement). This gives
a value of $k_{\text{flute}}=23.61$ and $k_{\text{kink}}=4.57$ for both
viscosity models. Hence, the observed most unstable fluting perturbation is
that of the form~\eqref{eq:kink_perturbation} where $m=4$ and $k=23.61$ and the
observed kink instability is that where $m=1$ and $k=4.57$. Using these values,
it is observed that the fluting perturbation exactly resonates with the field,
that is $m B_{\theta}(r)/r + kB_z(r) = 0$, at $r=0.125$
(Figure~\ref{fig:resonant_surface_4}). This is precisely the predicted radius
of peak linear \rev{flute growth $r_s$} seen in
Figure~\ref{fig:growth_rate_4}. At this time the perturbation is close
to resonance, that is $m B_{\theta}(r)/r + kB_z(r)\approx 0$, over a
range of radii from $r=0.125$ to $0.2$.

@@

Table~\ref{tab:kink_fluting_params} summarises the quantitative differences
between the results for the two values of the resistivity $\eta$. All values
are calculated from simulations using the anisotropic model with the exception
of $k_{\text{kink}}$ which is measured from isotropic results due to noise in
the anisotropic case (the value of $k_{\text{kink}}$ appears similar, however).
The results of the isotropic cases are qualitatively similar. The radius of
peak \rev{linear flute growth rate $r_s$} is calculated at time $t=20$. The fluting wavenumber $k$ and
observed $r_s$ are measured at times just prior to the nonlinear development of
the flute instability, that is at $t=22$ when $\eta=10^{-3}$ and $t=26$ when
$\eta = 10^{-4}$. The kink wavenumber is measured at $t=26$ in both cases.
These times allow fair comparison between measurements.

The longitudinal wavenumber $k_{\text{kink}}$ of the observed kink perturbation
remains similar in both cases since the instability is essentially governed by
the twist injected by the driver which remains the same in both cases. In
contrast, the longitudinal wavenumber $k_{\text{flute}}$ of the observed
fluting perturbation is lower in the $\eta=10^{-3}$ cases. This is due to the
different resonant surface within which the perturbation grows, the location
being dictated by the peak of the linear \ref{flute growth rate $r_s$}. Note that the location of
this peak again matches well the location of the observed resonant surface, as
in the $\eta=10^{-4}$ cases. Similar to the $\eta=10^{-4}$ cases, the \rev{peak
growth rate $r_s$} predicted by the linear analysis is the same order of magnitude as
the observed growth rate.
